---
title: "Visualisation Report"
output: html_document
date: "`r Sys.Date()`"
params:
  base_path:
    value: x
  sample_rate:
    value: x
  n_samples:
    value: 100
  n_col:
    value: 4
  window_length:
    value: 1
  overlap_percent:
    value: 0
  minimum_distance:
    value: 0.1
  shape_metric:
    value: "euclidean"
  num_neighbours:
    value: 10
  samples_to_analyse:
    value: 100
  
---

<!-- Unlike the data visualisation pipeline that Ryley is working on, this is a specific visualisation workflow for the koala data. This script will not generalise to other datasets. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 8, fig.height = 4) # smaller plots

library(data.table)
library(purrr)
library(tidyverse)
library(kableExtra)
```

```{r load data, include= FALSE}
# source functions from another file
source(file.path(params$base_path, "Scripts", "VisualiseBehaviours_New.R"))

# load in the raw data and cleaned feature data (e.g., remove redundant and NA features)
raw_data <- fread(file.path(params$base_path, "Data", "RawOtherData.csv"))
feature_data <- fread(file.path(params$base_path, "Data", "FeatureOtherData.csv"))
```

### Original data
This is the data as I received it. The classes are highly imbalanced which can lead to skewed performance.
```{r plot data volume, echo = FALSE}
activity_by_id <- plotActivityByID(data = raw_data, frequency = params$sample_rate)
activity_by_id$plot
```

Additionally, we should look at the quality. This can be assessed by looking at examples of each of the traces.

```{r generate trace plots, echo = FALSE}
tracePlots <- plotTraceExamples(
  behaviours = unique(raw_data$Activity),
  data = raw_data,
  individuals = length(unique(raw_data$ID)),
  n_samples = as.numeric(params$n_samples),
  n_col = as.numeric(params$n_col)
)
tracePlots$grid_plot
```

From this I can clean it up a bit by removing the meaning categories (blank and NA) then generate a small number of features from a subset of the data. We then use this feature data to generate a basic UMAP that could suggest clustering relationships between classes.

```{r process a sample of features, include = FALSE}
if(file.exists(file.path(params$base_path, "Data", "SampleFeatureData.csv"))){
  sample_feature_data <- fread(file.path(params$base_path, "Data", "SampleFeatureData.csv"))
} else {
  sample_data <- raw_data %>% 
    filter(!Activity %in% c("NaN", "")) %>% 
    group_by(Activity, ID) %>%
    slice(1:params$samples_to_analyse) %>% as.data.table()
  
  sample_feature_data <- generateFeatures(
    window_length = as.numeric(params$window_length), 
    sample_rate = as.numeric(params$sample_rate), 
    overlap_percent = as.numeric(params$overlap_percent), 
    raw_data = sample_data, 
    features_type = c("statistical", "timeseries"))
  
  fwrite(sample_feature_data, file.path(params$base_path, "Data", "SampleFeatureData.csv"))
}

sample_feature_cols <- removeBadFeatures(sample_feature_data, var_threshold = 0.5, cor_threshold = 0.9) 
selected_feature_data <- sample_feature_data %>%
  select(c(!!!syms(sample_feature_cols), "Activity")) %>% 
  na.omit()
```

```{r UMAP Visualisation, echo = FALSE}
numeric_features <- selected_feature_data %>% select(-"Activity")
labels <- selected_feature_data %>% select("Activity")

UMAPVis <- plotUMAPVisualisation(numeric_features, labels, minimum_distance = params$minimum_distance, num_neighbours = params$num_neighbours, shape_metric = params$shape_metric, spread = 2)
UMAPVis$UMAP_2D_plot
```

### Grouping the behaviours
Because in Gabby's research project she is only looking for some specific behaviours, I can simplify my analysis and improve performance by grouping stuff together into generalised classes. The classes I want to try to find are 'Walking', 'Still', 'Foraging/Eating' and 'Tree Movement' (with this last class being a generalised assortment). Examining the UMAP above, I can see some general trends that help me clump the behaviours together. 

For this analysis I have selected the following generalisations.

```{r generalisations}

behaviours_generalisations <- data.table(
  Activity = c(
    "Branch Walking", "Tree Movement", "Swinging/Hanging",
    "Climbing up", "Rapid Climbing", "Climbing Down",
    "Foraging/Eating", "Tree Sitting", "Ground Sitting",
    "Sleeping/Resting", "Shake", "Grooming",
    "Walking", "Bound/Half-Bound"
  )
)

# Add GeneralisedActivity column based on conditions
behaviours_generalisations[, GeneralisedActivity := fifelse(
  Activity %in% c("Branch Walking", "Tree Movement", "Swinging/Hanging"), "TreeMovement",
  fifelse(Activity %in% c("Climbing up", "Rapid Climbing", "Climbing Down"), "Climbing",
          fifelse(Activity %in% c("Foraging/Eating"), "Feed",
                  fifelse(Activity == "Tree Sitting", "TreeSitting",
                          fifelse(Activity == "Ground Sitting", "GroundSitting",
                                  fifelse(Activity %in% c("Sleeping/Resting"), "Still",
                                          fifelse(Activity %in% c("Shake", "Grooming"), "Groom",
                                                  fifelse(Activity %in% c("Walking", "Bound/Half-Bound"), "Walking", "NA")
                                          )))))))]

kable(data, caption = "Generalised functional groups for original activities")

```





